Improving Haskell Types with SMT

### 2015-06-05 docs/paper.tex
Correct type signatures.

To check for consistency, we assert all constraints, and ask the SMT
solver if the result is satisfiable.  If this is not the case, then
we report an error, otherwise we proceed with the algorithm.
\begin{Verbatim}
checkConsistent :: Solver -> [Expr] -> IO ()
checkConsistent s cs =
  do mapM_ (solverAssert s) cs
     res <- solverCheck s
     return (res == Sat)
\end{Verbatim}

@@

Unfortunately, not all SMT solvers support this feature and CVC4---%
the SMT solver that we used during the development of this work---%
does not have this ability.  To work around this, we implemented
a custom minimization algorithm which work as follows:
\begin{Verbatim}
minimize :: Solver ->
            [Expr] -> -- part of conflict
            [Expr] -> -- search for conflict
            IO ()
minimize s yes [] = return yes
minimize s yes mb =
  do solverPush s
     search s yes [] mb

minimize :: Solver ->
            [Expr] -> -- part of conflict
            [Expr] -> -- currently asserted
            [Expr] -> -- search for conflict
            IO ()
search s yes mb (c : cs) =
  do solverAssert s c
     res <- solverCheck s
     case res of
       Unsat ->
         do solverPop s
            solverAssert s c
            minimize s (c : yes) mb

       _ -> search s yes (c : mb) cs
\end{Verbatim}
The algorithm keeps track of constraints that we are certain are part
of the conflict in the variable \Verb"yes".  In addition, the algorithm
has the invariant that the constraints in \Verb"yes" are always asserted
in the solver's state, represented by the variable \Verb"s".
Initially, \Verb"yes" starts off empty.

>>>>

To check for consistency, we assert all constraints, and ask the SMT
solver if the result is satisfiable.  If this is not the case, then
we report an error, otherwise we proceed with the algorithm.
\begin{Verbatim}
checkConsistent :: Solver -> [Expr] -> IO Bool
checkConsistent s cs =
  do mapM_ (solverAssert s) cs
     res <- solverCheck s
     return (res == Sat)
\end{Verbatim}

@@

Unfortunately, not all SMT solvers support this feature and CVC4---%
the SMT solver that we used during the development of this work---%
does not have this ability.  To work around this, we implemented
a custom minimization algorithm which work as follows:
\begin{Verbatim}
minimize :: Solver ->
            [Expr] -> -- part of conflict
            [Expr] -> -- search for conflict
            IO [Expr]
minimize s yes [] = return yes
minimize s yes mb =
  do solverPush s
     search s yes [] mb

search :: Solver ->
          [Expr] -> -- part of conflict
          [Expr] -> -- currently asserted
          [Expr] -> -- search for conflict
          IO [Expr]

search s yes mb (c : cs) =
  do solverAssert s c
     res <- solverCheck s
     case res of
       Unsat ->
         do solverPop s
            solverAssert s c
            minimize s (c : yes) mb

       _ -> search s yes (c : mb) cs

search s yes mb [] = error "Impossible!"
\end{Verbatim}
The algorithm keeps track of constraints that we are certain are part
of the conflict in the variable \Verb"yes".  In addition, the algorithm
has the invariant that the constraints in \Verb"yes" are always asserted
in the solver's state, represented by the variable \Verb"s".
Initially, \Verb"yes" starts off empty.
